Quantifying stochastic noise in cultured\\ circadian reporter cells

### 2015-05-20 response_letter.tex

\begin{document}

\begin{flushright}
\today\\[2ex]
{\itshape Francis J. Doyle III}\\
Dept.\ of Chemical Engineering\\
Univ.\ of California, Santa Barbara\\
Santa Barbara, CA 93106-5080\\
\end{flushright}

{\itshape Christopher V. Rao}\\
Associate Editor\\
PLOS Computational Biology\\

{\itshape Douglas Lauffenburger}\\
Deputy Editor\\
PLOS Computational Biology\\

@@

We thank the reviewer for their positive critique of the study, and are excited that they would consider applying the method to their own work.

\begin{reviewer}
Main concerns:

M1. The author's comments notwithstanding, there are some concerns regarding the possible contribution of period heterogeneity to the quantified population damping. Specifically, the decays in Fig 3A actually looks somewhat like the decay in theoretical Figure S6, right panel, which was computed assuming period heterogeneity. Of course, the authors convincingly show that population period heterogeneity is less than peak-to-peak variation in percent. I still request a control experiment, in which simulated time-series of observed peak-to-peak variation AND observed population period heterogeneity are analyzed with the authors' algorithm. The method should be shown to work as for instance noise levels are increased or decreased also for this case. This could be a complement to Figure 1C, where the same thing was done for oscillators with identical period.
\end{reviewer}

These concerns were also shared by the first reviewer, and we have therefore {\bfseries substantially changed the language of the study to reflect the fact that extrinsic noise (period heterogeneity) could play a role in the damping rates} (see page 3 of this response).
For Fig. 3A, these profiles were obtained from a mathematical model of homogeneous cells.
The deviation from the exponential window is likely due to the limit cycle not being a perfect sinusoid.

@@

% I'll add a paragraph on practical considerations for implementing the algorithm. Sampling rate is less important than overall recording duration. There are many ways to do the nonlinear regression and analysis, which likely depends on the data being analyzed. For looking specifically at the damping rate, the best method is likely detrending, followed by a semilog plot of window vs.\ time, from which more traditional statistical techniques on comparing two regressions (ancova, bayesian inference)

\begin{reviewer}
Minor concerns:

Relating to question M1, the authors refer to their previous work, ref. 23, where it was shown that phase drift leads to an exponentially decreasing population signal. Otoh, period heterogeneity would lead to a time-squared exponent; where is the latter formula derived? If this was done by Rougemont and Naef, they should be cited.
\end{reviewer}

{\bfseries We have largely re-worded this section of the text} in response to the comments of other reviewers (see page 3 of this response). Despite this, we have added a further citation to Rougement \& Naef, as they present a more detailed theoretical analysis of the windowing function:

@@

We also note that the importance of amplitude boosting is linked to the measurement of noise, as noise likely plays a key role in determining the rate at which rhythms damp in peripheral tissues.

\begin{reviewer}
For the discussion: authors argue that the amplitude/noise ratio seem to have evolved to something close to an optimum. Yet, in the paper by Liu et al. in Cell 129, 2007, Cry2 knockouts seemed to have even stronger rhythms?
\end{reviewer}

This is an interesting point. We are careful in saying that there may indeed be perturbations or knockouts that shift the clock to a region of higher stability, but that on average perturbations tend to lead to less robust rhythms.

We unfortunately do not have access to the raw data of Liu et al.\ to look more closely at their observations. In the genome-wide siRNA screen, however, Cry2 knockdown was one of the several positive controls run in each plate.
From this data, it seems that Cry2 knockdown seems to simultaneously reduce both the population-level amplitude and damping rate:

### 2015-06-30 response_letter.tex

\begin{document}

\begin{flushright}
\today\\[2ex]
{\itshape Francis J. Doyle III}\\
Dept.\ of Chemical Engineering\\
Univ.\ of California, Santa Barbara\\
Santa Barbara, CA 93106-5080\\
\end{flushright}

{\itshape Christopher V. Rao}\\
Associate Editor\\
PLOS Computational Biology\\

{\itshape Douglas Lauffenburger}\\
Deputy Editor\\
PLOS Computational Biology\\

Thank you for your email on June 23th, 2015, inviting us to submit a further revised version of our manuscript, ``Quantifying stochastic noise in cultured circadian reporter cells'' (MS: PCOMPBIOL-D-15-00431) by Peter C. St. John and Francis J. Doyle~III.

We are pleased to hear that two of the reviewers recommend our work for publication. We thank the reviewers for their second round of thoughtful critiques, which we feel further improves the quality of the manuscript.

@@

Please see our detailed responses to the reviewers below:

\section*{Reviewer \#1}
\begin{reviewer}
The revision improved the manuscript, but a couple of minor issues remain to be addressed.

The added discussion on types of noise should be revised to include the work of Elowitz et al (Science 2002) on extrinsic vs intrinsic noise. The Elowitz study does quantitatively distinguish extrinsic vs intrinsic noise in their experiments, while the revised manuscript states this is only possible in theory (so the wording should be altered at the bottom of page 6 and the Elowitz study cited). 
\end{reviewer}

\begin{reviewer}
Also, in the revised part of the introduction addressing the role of noise in characterizing oscillations, the work by Bratsun et al (PNAS 2005) should be mentioned as another caution. For instance, they point out that ``the 'noise signature' of a system dominated by delay can be similar to a system dominated by intrinsic noise. This result implies that care must taken in attributing variability to purely stochastic sources, because delay-induced variability can appear empirically similar to fluctuations arising from intrinsic noise.''
\end{reviewer}

\begin{reviewer}
In the new expanded paragraph discussing the effect of cell period heterogeneity, the authors dismiss the significance of intrinsic differences in period across the population compared to cycle-to-cycle fluctuations and don’t directly quantify the relative contributions. However, Cohen et al (Journal Theoretical Biology 2013) quantified it for this same set of fibroblasts using a Bayesian hierarchical model and found the standard deviation of periods across the population to be 0.89h compared to the within-cell variability 1.43h. So while cycle-to-cycle fluctuations are indeed larger, the variability across the population is still quite substantial, and these values should be mentioned in this section to quantify the relative contributions of these sources of variability to the overall decay rate.
\end{reviewer}

@@

\section*{Reviewer \#4}

\begin{reviewer}
In short, the authors didn't really address the concern about applying their simple model for deductive reasoning. They claim that as the siRNA method is well-accepted by the circadian community that their interpretation should be legit. Biological response is proportional to dose and target. It seems implicitly that the authors are assuming that the DOSE of siRNA transfected into every individual cell is exactly the same. The biological response is then interpreted solely as being attributed to the target. One look at a flow cytometry result of a siRNA experiment will tell you that the dose is not the same between cells. Different siRNA also don't work equally well either. So again the criticism that the approach is not convincing in providing the claimed insight still is unaddressed.
\end{reviewer}

### 2015-07-02 response_letter.tex

{\itshape Christopher V. Rao}\\
Associate Editor\\
PLOS Computational Biology\\

{\itshape Douglas Lauffenburger}\\
Deputy Editor\\
PLOS Computational Biology\\

Thank you for your email on June 23th, 2015, inviting us to submit a further revised version of our manuscript, ``Quantifying stochastic noise in cultured circadian reporter cells'' (MS: PCOMPBIOL-D-15-00431) by Peter C. St. John and Francis J. Doyle~III.

We are pleased to hear that two of the four reviewers recommend our work for publication in {\itshape PLoS Computational Biology}. We thank the reviewers for their second round of thoughtful critiques, which we feel further improves the quality of the manuscript.

@@

\section*{Reviewer \#1}
\begin{reviewer}
The revision improved the manuscript, but a couple of minor issues remain to be addressed.

The added discussion on types of noise should be revised to include the work of Elowitz et al (Science 2002) on extrinsic vs intrinsic noise. The Elowitz study does quantitatively distinguish extrinsic vs intrinsic noise in their experiments, while the revised manuscript states this is only possible in theory (so the wording should be altered at the bottom of page 6 and the Elowitz study cited). 
\end{reviewer}

\begin{reviewer}
Also, in the revised part of the introduction addressing the role of noise in characterizing oscillations, the work by Bratsun et al (PNAS 2005) should be mentioned as another caution. For instance, they point out that ``the 'noise signature' of a system dominated by delay can be similar to a system dominated by intrinsic noise. This result implies that care must taken in attributing variability to purely stochastic sources, because delay-induced variability can appear empirically similar to fluctuations arising from intrinsic noise.''
\end{reviewer}

\begin{reviewer}
In the new expanded paragraph discussing the effect of cell period heterogeneity, the authors dismiss the significance of intrinsic differences in period across the population compared to cycle-to-cycle fluctuations and don't directly quantify the relative contributions. However, Cohen et al (Journal Theoretical Biology 2013) quantified it for this same set of fibroblasts using a Bayesian hierarchical model and found the standard deviation of periods across the population to be 0.89h compared to the within-cell variability 1.43h. So while cycle-to-cycle fluctuations are indeed larger, the variability across the population is still quite substantial, and these values should be mentioned in this section to quantify the relative contributions of these sources of variability to the overall decay rate.
\end{reviewer}

Indeed, we agree with the reviewer that adding quantitative measures adds to the discussion of the relative importance of intrinsic vs.\ extrinsic period variability. We have therefore added these values from both studies. In addition, we have softened some of the language to avoid dismissing the effect of extrinsic variability, which likely indeed plays a non-negligible role in determining the damping rate.

\begin{manuscript}{Page 6}
  While it is true that the dephasing of a group of oscillators can be caused by both variance in the mean period as well as cycle-to-cycle variability, intrinsic stochastic noise {\bf may} play a more significant role. 
{\bf We} show that there is greater variance in period on a cycle-to-cycle basis than on a cell-to-cell basis in cultured fibroblast cells {\bf (Figure S5): individual cell period lengths have an average inner quartile range (IQR) of 3.18 h, while cell-to-cell average period has an IQR of 1.55 h. 
These results are independently confirmed by a previous study using Bayesian modeling, which found a standard deviation of periods within cells of 1.43 h, and 0.89 h across the population (25). 
A similar} result has also been observed in dispersed SCN neurons, suggesting that while both intrinsic and extrinsic period heterogeneity likely contribute to the dephasing kinetics, cell-to-cell differences are less severe than cell-autonomous noise (2).

{\bf It is also} possible that damping rate changes due to siRNA or small molecule perturbation could be manifested through altering the system's extrinsic noise.

\end{manuscript}

### 2015-07-02 main_text.tex
second_revison
addressing point on including qualitative measures

\begin{figure}[tbp]
  \begin{center}
    % \includegraphics[width=0.8\textwidth]{figures/pdfs/Fig1.pdf}
  \end{center}
  \caption{{\bfseries Single-cell bioluminescence recordings show that higher stochastic noise results in faster damping at the population-level.}
  Data on the bioluminescence of single-cell fibroblasts was taken from Leise {\itshape et al.,} 2012 \cite{Leise2012}.
({\bfseries A}) Cells were sorted into two groups depending on their degree of stochastic noise. An example trace from each of the two groups is shown, demonstrating different levels of noise present in the dataset.
({\bfseries B}) After artificially synchronizing each cell, we calculate averaged bioluminescence rhythms of each group (solid lines). A damped sinusoid fit to both groups reveals a difference in damping rate, demonstrated by fitted envelope functions ($\pm A\exp{-dt}$, dashed lines).
({\bfseries C}) The observed absolute difference in damping rate was shown to be significant ($p = 0.0264$) using $10,000$ bootstrap trials.}
\label{fig:fibroblast_noise}
\end{figure}

@@

\begin{figure}[tbp]
  \begin{center}
    % \makebox[\textwidth][c]{\includegraphics[]{figures/pdfs/Fig2.pdf}}
  \end{center}
  \caption{{\bfseries Small molecule modulator KL001 increases damping rate in a dose-dependent fashion.}
  Experimental data on the dose-dependent effects of small molecules KL001 and longdaysin on cultured circadian reporter cells was taken from Hirota {\itshape et al.,} 2012 \cite{Hirota2012}.
({\bfseries A}) Detrended bioluminescence signals from the two reporter systems and two small molecules are shown normalized by the fitted amplitude, period, and phase. The normalized bioluminescence highlights the dose-dependent change in damping rate seen with the KL001 application (top), but not with longdaysin (bottom).
({\bfseries B}) Quantification of the dose-dependent change in damping rate caused by small molecule modulators. While both molecules lead to a dose-dependent increase in period, only KL001 shows a reliable change in damping rate.}
\label{fig:dose_dependence}
\end{figure}

@@

\begin{figure}[tbp]
  \begin{center}
    % \makebox[\textwidth][c]{\includegraphics[]{figures/pdfs/Fig3.pdf}}
  \end{center}
  \caption{{\bfseries Mathematical model accurately predicts dose-dependent changes in damping rate.}
({\bfseries A}) Example single-cell trajectories (top) and population-averaged trajectories (bottom, mean of $1,000$ cells) of cells under various treatments. Cells with the nominal parameter set (left, black) closely match the experimental damping rate for unperturbed cells. Cells with simulated KL001 action (red, center) are noisier at the single-cell level, and show faster damping at the population-level. Cells with simulated longdaysin action show slightly more accurate single-cell rhythms, with a corresponding reduction in the population-level damping rate.
({\bfseries B}) The model accurately predicts the general trend of period vs.\ damping rate for both KL001 and longdaysin perturbations. Experimental data points represent the mean of two replications at each concentration. Computational data points represent the mean of ten independent population simulations.}
\label{fig:simulation}
\end{figure}

@@

\begin{figure}[tbp]
  \begin{center}
    % \includegraphics[]{figures/pdfs/Fig4.pdf}
  \end{center}
  \caption{{\bfseries Fit quality vs.\ damping rate for the genome-wide siRNA screen.}
({\bfseries A}) A plot of the $111,743$ individual fits shows that the majority of fits have a high $R^2$ value and positive damping rate. Only fits with $R^2 > 0.8$ were kept for further analysis.
({\bfseries B}) Examples chosen randomly from each of the four quadrant regions in (A). Sinusoidal parameters for fits 1-2 can be more confidently inferred than those for fits 3-4.}
\label{fig:fit_quality}
\end{figure}

@@

\begin{figure}[tbp]
  \begin{center}
    % \includegraphics[width=0.8\textwidth]{figures/pdfs/Fig5.pdf}
  \end{center}
  \caption{{\bfseries Distributions in fitted parameters for the genome-wide siRNA screen.} For each well in the high-throughput screen, the period, amplitude and damping rate are calculated. After normalization, distributions in robust z-scores closely resemble normal distributions. For all parameters, the region of highest density is consistent between the control and perturbed populations, indicating many perturbations do not appreciably change clock dynamics.} 
\label{fig:fit_distributions}
\end{figure}

\begin{table}
  \begin{center}
    \begin{tabular}{lrrrrrr}\toprule
      {} & \multicolumn{2}{c}{$T$} & \multicolumn{2}{c}{$\ln A$} & \multicolumn{2}{c}{$d$} \\
      {}         & \multicolumn{1}{c}{C}         & \multicolumn{1}{c}{P}           & \multicolumn{1}{c}{C}         & \multicolumn{1}{c}{P}               & \multicolumn{1}{c}{C}         & \multicolumn{1}{c}{P}           \\\midrule
      Mean     & $-0.234$ & $0.187$ & $0.443$  & $-0.343$ & $0.043$  & $0.090$    \\
      Variance & $ 0.599$ & $3.313$ & $0.605$  & $ 3.074$ & $0.771$  & $2.850$    \\
      Skewness & $ 0.153$ & $0.367$ & $-1.823$ & $-0.580$ & $-0.107$ & $0.371$    \\
      Kurtosis & $ 3.772$ & $0.591$ & $8.329$  & $ 0.476$ & $2.423$  & $0.373$    \\
      \bottomrule
    \end{tabular}
  \end{center}
  \caption{{\bfseries Moments of the fitted parameter distributions (Figure~\ref{fig:fit_distributions}) after normalization and outlier removal.} Parameters were normalized by subtracting the median and dividing by the median absolute deviation on a plate-by-plate basis. Control distributions had less variance, were less skewed, and were more peaked then their perturbed counterparts.}
  \label{tab:fit_distributions}
\end{table}

@@

\begin{table}
  \begin{center}
    \begin{tabular}{rrrrr}
      \toprule
      {}       & $d$    & $\ln A$ & $T$    & $\theta$ \\\midrule
      $d$      & $1.000 $ & $0.285 $  & $-0.142$ & $-0.269$\\
      $\ln A$  & $0.285 $ & $1.000 $  & $-0.022$ & $-0.112$\\
      $T$      & $-0.142$ & $-0.022$  & $1.000 $ & $-0.113$\\
      $\theta$ & $-0.269$ & $-0.112$  & $-0.113$ & $1.000 $\\
      \bottomrule
    \end{tabular}
  \end{center}
  \caption{{\bfseries Correlation among normalized parameters of the high-throughput siRNA screen.} Pearson correlation coefficients are relatively low between fitted parameters, indicating that changes to damping rate (and thereby stochastic noise) are not explained by changes to period, amplitude, or phase.}
  \label{tab:corr}
\end{table}

@@

\begin{figure}[tbp]
  \begin{center}
    % \includegraphics[width=0.65\textwidth]{figures/pdfs/Fig6.pdf}
  \end{center}
  \caption{{\bfseries Effects of siRNA knockdowns on amplitude and damping rate.}
  Clock robustness is a function of both amplitude and damping rate.
  Distributions in amplitude and damping rate for control wells ({\bfseries A}) or perturbed wells ({\bfseries B}) indicate that perturbations tend to shift the clock towards regions of higher damping rate or lower amplitude. Green dots in each figure indicates the mean of the control population.
({\bfseries C}) Averaged effect of siRNA knockdown after grouping the perturbed population by Gene ID. Only those genes that were significantly different from the control distribution are shown (Hotelling's $T^2$ test, $\alpha = 0.01$).
({\bfseries D}) Radial histogram of the significant gene perturbations shown in (C). The area of each slice is proportional to the frequency of perturbations away from the mean in that direction. Very few gene knockdowns result in both higher amplitude and lower damping rate (red slices, lower right quadrant).
}
\label{fig:outlier_dist}
\end{figure}

@@

In the preceding sections, we have demonstrated that changes to single-cell intrinsic noise are sufficient to explain the observed changes in population-level damping.
However, experimental work has shown that cell-autonomous fibroblast cells have a distribution of mean free-running periods \cite{Leise2012}.
Indeed, prior to the availability of single-cell data, studies explored the possibility that differences in mean periods served as the mechanism behind population-level damping \cite{Izumo2003}.
While it is true that the dephasing of a group of oscillators can be caused by both variance in the mean period as well as cycle-to-cycle variability, intrinsic stochastic noise likely plays a more significant role: we show that there is greater variance in period on a cycle-to-cycle basis than on a cell-to-cell basis in cultured fibroblast cells (Figure~\ref{fig:cell2cell-variability}).
This result has also been observed in dispersed SCN neurons, indicating cell heterogeneity is less severe than cell-autonomous noise \cite{Herzog2004}.

However, it is possible that damping rate changes due to siRNA or small molecule perturbation could be manifested through altering the system's extrinsic noise.
Such a change, caused perhaps by an unequal uptake of drug across cells, would result increased cell heterogeneity, and different dephasing kinetics.
While differentiating between intrinsic and extrinsic noise sources from population-level data is possible in theory (Figure~\ref{fig:dephasing-rates}), these differences are likely not identifiable from typical data (Figure~\ref{fig:computational_dephasing}).
Differences in damping rates are therefore best viewed as representative of changes to overall stochastic noise from both intrinsic and extrinsic factors.
Since both types of noise are important to determining the overall function of population-level rhythms, damping rates are still a valuable method of quantifying stochastic noise.

@@

\section*{Acknowledgments}
We thank Tsuyoshi Hirota, David Welsh, and John Abel for the helpful discussions, as well as the each of the authors of the original experiments for making their raw data available online.
This work was supported by the National Institutes of Health/National Institute of General Medical Sciences under award number 1R01GM096873-01 and by the Institute for Collaborative Biotechnologies through grant W911NF-09-0001 from the U.S.\ Army Research Office.

@@

>>>>

@@

@@

@@

@@

@@

@@

@@

In the preceding sections, we have demonstrated that changes to single-cell intrinsic noise are sufficient to explain the observed changes in population-level damping.
However, experimental work has shown that cell-autonomous fibroblast cells have a distribution of mean free-running periods \cite{Leise2012}.
Indeed, prior to the availability of single-cell data, studies explored the possibility that differences in mean periods served as the mechanism behind population-level damping \cite{Izumo2003}.
While it is true that the dephasing of a group of oscillators can be caused by both variance in the mean period as well as cycle-to-cycle variability, intrinsic stochastic noise may play a more significant role. 
We show that there is greater variance in period on a cycle-to-cycle basis than on a cell-to-cell basis in cultured fibroblast cells (Figure~\ref{fig:cell2cell-variability}): individual cell period lengths have an average inner quartile range (IQR) of 3.18 h, while cell-to-cell average period has an IQR of 1.55 h. 
These results are independently confirmed by a previous study using Bayesian modeling, which found a standard deviation of periods within cells of 1.43 h, and 0.89 h across the population \cite{Cohen2012}.
A similar result has also been observed in dispersed SCN neurons, suggesting that while both intrinsic and extrinsic period heterogeneity likely contribute to the dephasing kinetics, cell-to-cell differences are less severe than cell-autonomous noise \cite{Herzog2004}.

It is also possible that damping rate changes due to siRNA or small molecule perturbation could be manifested through altering the system's extrinsic noise.
Such a change, caused perhaps by an unequal uptake of drug across cells, would result increased cell heterogeneity, and different dephasing kinetics.
While differentiating between intrinsic and extrinsic noise sources from population-level data is possible in theory (Figure~\ref{fig:dephasing-rates}), these differences are likely not identifiable from typical data (Figure~\ref{fig:computational_dephasing}).
Differences in damping rates are therefore best viewed as representative of changes to overall stochastic noise from both intrinsic and extrinsic factors.
Since both types of noise are important to determining the overall function of population-level rhythms, damping rates are still a valuable method of quantifying stochastic noise.

@@

\section*{Acknowledgments}
We thank Tsuyoshi Hirota, David Welsh, and John Abel for the helpful discussions, as well as the each of the authors of the original experiments for making their raw data available online.
% This work was supported by the National Institutes of Health/National Institute of General Medical Sciences under award number 1R01GM096873-01 and by the Institute for Collaborative Biotechnologies through grant W911NF-09-0001 from the U.S.\ Army Research Office.

@@

\clearpage

\begin{figure}[h!]
  \begin{center}
    % \includegraphics[width=0.8\textwidth]{figures/pdfs/Fig1.pdf}
  \end{center}
  \caption{{\bfseries Single-cell bioluminescence recordings show that higher stochastic noise results in faster damping at the population-level.}
  Data on the bioluminescence of single-cell fibroblasts was taken from Leise {\itshape et al.,} 2012 \cite{Leise2012}.
({\bfseries A}) Cells were sorted into two groups depending on their degree of stochastic noise. An example trace from each of the two groups is shown, demonstrating different levels of noise present in the dataset.
({\bfseries B}) After artificially synchronizing each cell, we calculate averaged bioluminescence rhythms of each group (solid lines). A damped sinusoid fit to both groups reveals a difference in damping rate, demonstrated by fitted envelope functions ($\pm A\exp{-dt}$, dashed lines).
({\bfseries C}) The observed absolute difference in damping rate was shown to be significant ($p = 0.0264$) using $10,000$ bootstrap trials.}
\label{fig:fibroblast_noise}
\end{figure}

\begin{figure}[h!]
  \begin{center}
    % \makebox[\textwidth][c]{\includegraphics[]{figures/pdfs/Fig2.pdf}}
  \end{center}
  \caption{{\bfseries Small molecule modulator KL001 increases damping rate in a dose-dependent fashion.}
  Experimental data on the dose-dependent effects of small molecules KL001 and longdaysin on cultured circadian reporter cells was taken from Hirota {\itshape et al.,} 2012 \cite{Hirota2012}.
({\bfseries A}) Detrended bioluminescence signals from the two reporter systems and two small molecules are shown normalized by the fitted amplitude, period, and phase. The normalized bioluminescence highlights the dose-dependent change in damping rate seen with the KL001 application (top), but not with longdaysin (bottom).
({\bfseries B}) Quantification of the dose-dependent change in damping rate caused by small molecule modulators. While both molecules lead to a dose-dependent increase in period, only KL001 shows a reliable change in damping rate.}
\label{fig:dose_dependence}
\end{figure}

\begin{figure}[h!]
  \begin{center}
    % \makebox[\textwidth][c]{\includegraphics[]{figures/pdfs/Fig3.pdf}}
  \end{center}
  \caption{{\bfseries Mathematical model accurately predicts dose-dependent changes in damping rate.}
({\bfseries A}) Example single-cell trajectories (top) and population-averaged trajectories (bottom, mean of $1,000$ cells) of cells under various treatments. Cells with the nominal parameter set (left, black) closely match the experimental damping rate for unperturbed cells. Cells with simulated KL001 action (red, center) are noisier at the single-cell level, and show faster damping at the population-level. Cells with simulated longdaysin action show slightly more accurate single-cell rhythms, with a corresponding reduction in the population-level damping rate.
({\bfseries B}) The model accurately predicts the general trend of period vs.\ damping rate for both KL001 and longdaysin perturbations. Experimental data points represent the mean of two replications at each concentration. Computational data points represent the mean of ten independent population simulations.}
\label{fig:simulation}
\end{figure}

\begin{figure}[h!]
  \begin{center}
    % \includegraphics[]{figures/pdfs/Fig4.pdf}
  \end{center}
  \caption{{\bfseries Fit quality vs.\ damping rate for the genome-wide siRNA screen.}
({\bfseries A}) A plot of the $111,743$ individual fits shows that the majority of fits have a high $R^2$ value and positive damping rate. Only fits with $R^2 > 0.8$ were kept for further analysis.
({\bfseries B}) Examples chosen randomly from each of the four quadrant regions in (A). Sinusoidal parameters for fits 1-2 can be more confidently inferred than those for fits 3-4.}
\label{fig:fit_quality}
\end{figure}

\begin{figure}[h!]
  \begin{center}
    % \includegraphics[width=0.8\textwidth]{figures/pdfs/Fig5.pdf}
  \end{center}
  \caption{{\bfseries Distributions in fitted parameters for the genome-wide siRNA screen.} For each well in the high-throughput screen, the period, amplitude and damping rate are calculated. After normalization, distributions in robust z-scores closely resemble normal distributions. For all parameters, the region of highest density is consistent between the control and perturbed populations, indicating many perturbations do not appreciably change clock dynamics.} 
\label{fig:fit_distributions}
\end{figure}

\begin{figure}[h!]
  \begin{center}
    % \includegraphics[width=0.65\textwidth]{figures/pdfs/Fig6.pdf}
  \end{center}
  \caption{{\bfseries Effects of siRNA knockdowns on amplitude and damping rate.}
  Clock robustness is a function of both amplitude and damping rate.
  Distributions in amplitude and damping rate for control wells ({\bfseries A}) or perturbed wells ({\bfseries B}) indicate that perturbations tend to shift the clock towards regions of higher damping rate or lower amplitude. Green dots in each figure indicates the mean of the control population.
({\bfseries C}) Averaged effect of siRNA knockdown after grouping the perturbed population by Gene ID. Only those genes that were significantly different from the control distribution are shown (Hotelling's $T^2$ test, $\alpha = 0.01$).
({\bfseries D}) Radial histogram of the significant gene perturbations shown in (C). The area of each slice is proportional to the frequency of perturbations away from the mean in that direction. Very few gene knockdowns result in both higher amplitude and lower damping rate (red slices, lower right quadrant).
}
\label{fig:outlier_dist}
\end{figure}

\clearpage

\begin{table}[h!]
  \begin{center}
    \begin{tabular}{lrrrrrr}\toprule
      {} & \multicolumn{2}{c}{$T$} & \multicolumn{2}{c}{$\ln A$} & \multicolumn{2}{c}{$d$} \\
      {}         & \multicolumn{1}{c}{C}         & \multicolumn{1}{c}{P}           & \multicolumn{1}{c}{C}         & \multicolumn{1}{c}{P}               & \multicolumn{1}{c}{C}         & \multicolumn{1}{c}{P}           \\\midrule
      Mean     & $-0.234$ & $0.187$ & $0.443$  & $-0.343$ & $0.043$  & $0.090$    \\
      Variance & $ 0.599$ & $3.313$ & $0.605$  & $ 3.074$ & $0.771$  & $2.850$    \\
      Skewness & $ 0.153$ & $0.367$ & $-1.823$ & $-0.580$ & $-0.107$ & $0.371$    \\
      Kurtosis & $ 3.772$ & $0.591$ & $8.329$  & $ 0.476$ & $2.423$  & $0.373$    \\
      \bottomrule
    \end{tabular}
  \end{center}
  \caption{{\bfseries Moments of the fitted parameter distributions (Figure~\ref{fig:fit_distributions}) after normalization and outlier removal.} Parameters were normalized by subtracting the median and dividing by the median absolute deviation on a plate-by-plate basis. Control distributions had less variance, were less skewed, and were more peaked then their perturbed counterparts.}
  \label{tab:fit_distributions}
\end{table}

\begin{table}[h!]
  \begin{center}
    \begin{tabular}{rrrrr}
      \toprule
      {}       & $d$    & $\ln A$ & $T$    & $\theta$ \\\midrule
      $d$      & $1.000 $ & $0.285 $  & $-0.142$ & $-0.269$\\
      $\ln A$  & $0.285 $ & $1.000 $  & $-0.022$ & $-0.112$\\
      $T$      & $-0.142$ & $-0.022$  & $1.000 $ & $-0.113$\\
      $\theta$ & $-0.269$ & $-0.112$  & $-0.113$ & $1.000 $\\
      \bottomrule
    \end{tabular}
  \end{center}
  \caption{{\bfseries Correlation among normalized parameters of the high-throughput siRNA screen.} Pearson correlation coefficients are relatively low between fitted parameters, indicating that changes to damping rate (and thereby stochastic noise) are not explained by changes to period, amplitude, or phase.}
  \label{tab:corr}
\end{table}
